---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# halfsibdesign

<!-- badges: start -->
<!-- badges: end -->

A package containing several algorithms for finding REML estimates for covariances
in the balanced $q$-dimensional half-sib design
*y<sub>ijk</sub> = $\mu$ + $\alpha$<sub>i</sub> + $\beta$<sub>ij</sub> + $\epsilon$<sub>ijk</sub>*,
*$1\leq i \leq I, 1\leq j \leq J, 1\leq k \leq K$*,
where 
*$\alpha$<sub>i</sub>* $\sim \mathcal{N}(0, A)$,
*$\beta$<sub>ij</sub>* $\sim \mathcal{N}(0, B)$,
*$\epsilon$<sub>ijk</sub>* $\sim \mathcal{N}(0, E)$.

## Installation

You can install the development version of halfsibdesign from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("damian-t-p/halfsibdesign")
```

## Example

Simulate a half-sib experiment with specified parameters:

```{r example}
set.seed(1)

library(halfsibdesign)

q <- 4 # number of traits

I <- 100 # number of sires
J <- 3 # number of dams
K <- 5 # number of individuals per line

mu <- 1:q

sigma_a <- 5
sigma_b <- 3
sigma_e <- 1

A <- sigma_a^2 * diag(c(0, 0, 1, 1))
B <- sigma_b^2 * diag(q)
E <- sigma_e^2 * diag(q)

df <- rhalfsib(mu, A, I, B, J, E, K)
```

First, perform a MANOVA fit
```{r}
manova_2way(df)
```

Notice that `S3`, the within-sires estimate is not non-negative definite, se we 
compute the correponding REML estimate.

```{r}
stepreml_2way(df)
```


